The Talking Heads experiment

### 2015-04-29 chap10/chap10.tex
part 2 of martin haspelmath's comments

These experiments extended the Talking Heads experiment in several directions: 
\begin{enumerate}
\item {\itshape Vision.} The pan-tilt cameras and the flat white board with geometric figures ensured that the vision system 
of the Talking Heads was relatively reliable, although light conditions and camera misalignment
could occasionally 
cause havoc. The \textsc{aibo} experiments now used 3d objects in the real world that were seen from many different 
angles and in different light conditions (see \figref{f:look-ball}, right). This obviously pushed up the 
difficulty of visual perception enormously and it also made it much more difficult to acquire stable categories 
for reliably recognising the objects in the environment. 
\item {\itshape Speech.} The interaction with the robot used spoken language, which implied that the robot needed 
speech synthesis and a speech recognition system that could acquire new words. These components were available on the 
robot and could therefore be easily integrated. 
\item {\itshape Flexible scripts.} The Talking Heads experiments used a single very streamlined interaction scenario whereby
the robots played a Guessing Game with a clear turn-taking script. The \textsc{aibo} experiments targeted different 
kinds of interactions with humans and therefore the robot needed not only a way to represent and execute these 
different scripts but also a much more flexible way to move from one step to another within a script and to move 
between scripts. Moreover the scripts had to be integrated intimately with the physical behaviours that were 
steering the robot itself, which could in many cases not be overridden or directly controlled. 
\end{enumerate}

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap10/figs/look-ball.pdf}
\includegraphics[width=.45\textwidth]{chap10/figs/views-ball.pdf}}
\caption{\label{f:look-ball}Left: Fr\'{e}d\'{e}ric Kaplan interacts with the \textsc{aibo} robot in an experiment for the social learning of lexicons.
Right: View of the ball through the eye of the robot. As the same object is now seen from many different viewpoints it 
becomes much more challenging to consistently recognise it.}
\end{figure}
The first \textsc{aibo} experiments did not involve a population of agents but rather a single agent that was interacting directly with a human
experimenter. The agent had to learn the lexicon of the human through social 
interaction within the context of situated language games. The 2001
paper \dblenquote{\textsc{aibo}'s first words} by Steels and Kaplan \nocite{Steels:01h} described experiments in which an \textsc{aibo} robot 
learned words about objects in the environment, such as \smplenquote{ball} or \smplenquote{smiley}, as well as words for actions, such as 
\smplenquote{sit down} or \smplenquote{stand up}.

@@

The Perspective Reversal experiment \is{Perspective Reversal} focused on the question how a spatial language could emerge that employed
perspective reversal. It was the first experiment to systematically compare different language strategies, thus 
introducing for the first time the {\itshape comparative method} \is{comparative method} 
that would inform all later language game experiments.

A {\itshape language strategy} \is{language strategy} 
prescribes how some aspect of meaning needs to be conceptualised, expressed, and acquired. 
The Talking Heads experiment used a single strategy: find a discriminating category for the topic and employ a word 
for that category. But languages use a variety of strategies all intermixed. 
For example, speakers might have a strategy for expressing argument structure using cases (as in Latin or German) or 
a strategy for expressing tense and aspect using morphological markers on the verb (as in Spanish). Each strategy gives 
rise to a particular feature of a language and in order to demonstrate why languages exhibit this feature it is therefore
possible to compare a language with the feature to one without the feature by doing an experiment where
agents are endowed with a strategy that generates this feature and another one where they employ a different strategy
that does not. The performance of the resulting system, for example, its average communicative success, the time to reach 
linguistic convergence, the amount of cognitive effort involved, can be measured and thus the adaptive value of one 
strategy (and hence one feature of language) can be compared with another.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap10/figs/tokyo-persp.pdf}}
\caption{\label{fig:tokyo-persp}Luc Steels demonstrates live the Perspective Reversal experiment in Tokyo, April 2005. \textsc{aibo} robots track 
the movement of a ball and 
the robot acting as speaker describes to the hearer an observed movement, possibly taking into account the 
perspective of the hearer. For example, the speaker might say \smplenquote{the ball rolls from my left to your right}.}
\end{figure}

\enlargethispage{1\baselineskip}
The Perspective Reversal experiment worked as follows. A population of \textsc{aibo} robots roam around
freely in an unconstrained in-door environment. As soon as one sees the
ball, it comes to a stop and searches for another robot nearby, which also
looks for the ball and stops when it sees it. Then the human
experimenter pushes the ball with a stick so that it rolls a short
distance (see \figref{fig:tokyo-persp}). The two robots next play a description game. One robot (acting as the \smplenquote{speaker}) 
describes the ball-moving event to the other robot (the \dblenquote{hearer}), going through the same semiotic cycle as the Talking 
Heads agents. The game is a success if the meaning conveyed by the speaker is compatible with the hearer's 
own perception of the scene.

@@

Within this experimental setting three different strategies were tried in a long-term experiment where
a new strategy was injected at certain stages and performance monitored: 
\begin{itemize}
\item In stage \textsc{i}, agents use 
the Discrimination Game strategy and the Naming Game strategy discussed in chapters 4 and 5. A 
lexicon is emerging and there are some successful games (those for which the speaker and the hearer see the situation from 
the same perspective). But the lexicon keeps expanding and success is very limited as games where robots have a different 
perspective on the rolling ball all fail. 
\item In stage \textsc{ii}, agents have an extended 
strategy where agents can geometrically transform a model of the scene, based on their own perspective, into a model of the scene 
as seen from the perspective of the hearer. They can then conceptualise the scene from that perspective. 
Communicative success significantly increases and the 
lexicon starts to decrease. However there is ambiguity because the hearer cannot know whether
the speaker has conceptualised the 
scene from his own viewpoint or performed a perspective reversal, leading to a large number of failed communications 
(more than 50 \%). 
\item In stage \textsc{iii}, a more elaborate strategy is used. The speaker now expresses whether he has performed perspective 
reversal (as in \smplenquote{{\itshape to your} left} vs. simply \smplenquote{left}). Only with this strategy are agents able to reach high 
communicative success with a stable inventory. 
\end{itemize}

@@

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/associations.pdf}}
\caption{\label{fig:associations}Top: Two lexical associations have become active (association-259 and association-234) to cover the distinctive 
feature-set ([VISIB v-411] [VER v-431]) resulting from discrimination and perception. Bottom: 
The group of words and the meanings match with schemas forming part of a grammatical association that combines the words. }
\end{figure}
An application of these schemas in producing the utterance \smplenquote{(W O) (W U)} is represented in \figref{fig:associations} and shown 
here: 
\begin{verbatim}
127 ++> Speaker: head-40 Topic: 2 Context: 6 5 4 3 1 0
Categorial Perception:
  ([VISIB v-4111] [VER v-431])
  ([VER v-4311] [\textsc{area} v-4061])
Lexicon lookup: (Association-259 Association-234)
Syntactic structure:
  (syn-cat-77
    (syn-slot-50 (syn-cat-75 |(W O)|))
    (syn-slot-51 (syr-cat-76 |(W U)|)))
Semantic structure:
  (sem-cat-78
    (sem-slot-50 (sem-cat-75 (VER v-431)))
    (sem-slot-51 (sem-cat-76 (VISIB v-411))))
Meaning:
  ([VISIB v-4111] [VER v-4311)
Expression: ((W 0) (W U))
\end{verbatim}

@@

But first we need a clearer idea on what case grammar is.
In the sentence \smplenquote{John gave Mary a book}, John plays the role of agent, the giver 
of the book, Mary that of recipient or 
beneficiary and the book is the \dblenquote{patient} or object of the action. Languages differ in terms of the strategies that they 
use to express these participant roles. Some of them, like German, have an elaborate system of morphological 
markers attached to constituents of the noun phrase. These markers signal that the 
phrase is nominative, accusative, dative, etc., and these cases then map in complex ways 
onto semantic roles such as agent and patient.
Other languages, like English, use a strategy based on word order and prepositions. For example, 
\smplenquote{Mary gave John a book} contains the same participants as \smplenquote{John gave Mary a book} but now Mary is the 
agent and John the recipient. Still other languages, like Japanese, use a strategy based on particles following 
the noun, as in the Japanese sentence \smplenquote{Tanaka-san wa Tokyo de o-to-san ni atta} 
meaning Tanaka met his father in Tokyo. Tanaka-san is the name of a person, oto-san means \smplenquote{father} and atta 
means \smplenquote{met}. The markers are \smplenquote{wa} for the topic, \smplenquote{de} for location and \smplenquote{ni} for another sense of location which can 
also function as possessive. The case grammar experiments focused on modelling the latter strategy.

It used the comparative method again to compare three progressively more complex strategies: 
\begin{enumerate}
\item {\bfshape  Lexical strategy}. The experiment starts from a purely 
lexical language that is predefined in all agents. The lexicon contained words for expressing properties of 
objects (such as \smplenquote{smooth}, \smplenquote{red}, etc.)
as well as actions involving objects (such as \smplenquote{move-away-from}, \smplenquote{move-towards} or \smplenquote{push}). English words
were chosen to make the experiment easier to follow. Agents form
multi-word utterances using straightforward lexicon lookup, producing utterances
like \smplenquote{move-towards red jill} (Jill moves towards the red object) or `push 
jack jill green' (Jack pushes the green block to Jill). There is no syntax at all. Agents can nevertheless already 
achieve communicative success when they use the real world as a way to figure out the participant roles, even though the 
utterance does not provide any information. For example, to detect whether Jack pushes the green block to Jill or Jill 
pushes the green block to Jack.

\item {\bfshape  Specific marker strategy}. The second strategy introduces markers in the form of particles added to words referring 
to objects, in order to specify what argument the object fills in the predicate describing the 
event. For example, there could be two markers: \smplenquote{pi} 
and \smplenquote{pa} specifying that the referent plays the role respectively of giver and gift in a give event, as in
\smplenquote{pi jill pa block give}. The markers are entirely specific for each event in the sense that there is no generalisation 
yet to 'agent', 'patient' or other more abstract semantic roles. This already yields a more effective 
communication system because the complexity of seeking an interpretation matching with the world model is heavily
reduced and residual ambiguity (i.e. there is still more than one way in which the 
meaning matches with the world model) is also avoided.

@@

\subsubsection{Object identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and 'latches onto' regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or \dblenquote{anchors} that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\textsc{cogar} represents syntactic and semantic structures as frames with units, slots and 
constraints on the slots. The frames are in many ways similar to the feature structures used in feature-structure 
based grammars such as \textsc{hpsg}, although there are many important differences. For example, each unit has a 
unique name so that it becomes possible to refer back to units in defining constraints. \textsc{cogar} also used 
logic variables (denoted by a symbol with a question mark in front) as opposed to path descriptions, which 
lead to the adoption of techniques from logic programming. 
The syntactic and semantic structure of a particular utterance are coupled together as a coupled feature structure 
by making the unit-names shared. Often both are combined when the coupled feature structure needs to be displayed
as in this example below. 
\begin{verbatim}
((unit-15
   (scope utterance)
   (subunits (unit-17 unit-18 unit-16))
   (ordering (unit-17 unit-18 unit-16)))
 (unit-17
   (form (smooth))
   (meaning ((smooth ?source-object1)))
   (referent ?source-object1) 
   (gramcat (takes-marker-1)))
 (unit-18 (form (po)))
 (unit-16 
   (form (move-away-from)) 
   (meaning
     ((move-away-from ?event ?state) 
      (move-away-from-1 ?event ?object)
      (move-away-from-2 ?event ?source-object1)))))
\end{verbatim}
This example could have been constructed, either in parsing or producing, for 
the utterance \smplenquote{po smooth move-away-from}. Syntactic properties such as the form of the word, 
the ordering of the units, or the grammatical categories of a unit are represented explicitly on the syntactic side and 
the meaning and referent are represented on the semantic side. In general, \textsc{cogar} allows the representation of
any kind of information at whatever level of linguistic analysis as part of a feature structure.

@@

Here is an example of a lexical rule that defines the meaning of the word for an event \smplenquote{move-away-from}. The referent is 
the object that is referred to by the word, in this case it is the event itself. 
\clearpage
\begin{verbatim}
((?unit 
   (meaning
      (== (move-away-from ?event ?state)
          (move-away-from-1 ?event ?object)
          (move-away-from-2 ?event ?source)))
   (referent ?event)) 
<-->
((?unit
   (form (== move-away-from)))))
\end{verbatim}
Next, here is an example of a rule for the marker \smplenquote{po}. On the semantic side it ensures that the referent of one unit 
whose referent is ?source is equal to the object filling the argument move-away-from-2 in a move-away-from event. 
On the syntactic side, it triggers when there is a word that has the grammatical category \smplenquote{takes-marker-1} and 
is preceded by the marker \smplenquote{po}. (<< stands for \smplenquote{precedes}) 
\begin{verbatim}
((?predicate 
   (meaning
    (== (move-away-from ?evnt)
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source))))
 (?argument
   (referent ?source))) 
<---> 
((?argument
   (gramcat (== takes-marker-1)))
 (?marker-unit
   (form (== po))) 
(<< ?marker-unit ?argument))))
\end{verbatim}

There are additional rules that specify which word takes which marker, by assigning them to a particular 
grammatical category. For example, the word \smplenquote{smooth} has the category \smplenquote{takes-marker-1}: 
\begin{verbatim}
((?unit
   (gramcat (== takes-marker-1)) 
   (form (== smooth))))
<---> 
((?unit (form (== smooth))))
\end{verbatim}

The marker \smplenquote{po} in the above example is entirely specific for one argument of the move-away-from predicate, as would be the 
case when the second strategy is enacted. To implement the third strategy 
(to achieve more abstract semantic roles) requires that the markers become generic, which implies that 
we get additional rules that recategorise the arguments of a predicate in terms of 
semantic roles like: agent, patient, source, etc. Here is an example of a rule that 
has this effect. role-1 is an example of a semantic role created by the agent as an abstraction of the 
move-away-from-2 argument. 
\begin{verbatim}
((?predicate
   (meaning
    (== (move-away-from ?evnt) 
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source)))))
<---> 
 ((?predicate
    (meaning
      (== (move-away-from ?evnt)
          (move-away-from-1 ?evnt ?patient)
          (move-away-from-2 ?evnt ?source)))) 
    (expanded-meaning (== (role-1 ?evnt ?source))))
\end{verbatim}
And the rules that click everything together now operate over semantic roles as shown in the following example: 
\begin{verbatim}
 ((?predicate
   (expanded-meaning (== (role-1 ?event ?source))))
  (?unit (referent ?source))) 
<---> 
 ((?unit (gramcat (== takes-marker-1)))
  (?marker-unit (form (== po))) 
 (<< ?marker-unit ?unit))
\end{verbatim}
Recall that all rules must be reversible and that the conditional part must match and the concluding part must unify.

>>>>

These experiments extended the Talking Heads experiment in several directions: 
\begin{enumerate}
\item {\bfshape Vision.} The pan-tilt cameras and the flat white board with geometric figures ensured that the vision system 
of the Talking Heads was relatively reliable, although light conditions and camera misalignment
could occasionally 
cause havoc. The \textsc{aibo} experiments now used 3d objects in the real world that were seen from many different 
angles and in different light conditions (see \figref{f:look-ball}, right). This obviously pushed up the 
difficulty of visual perception enormously and it also made it much more difficult to acquire stable categories 
for reliably recognising the objects in the environment. 
\item {\bfshape Speech.} The interaction with the robot used spoken language, which implied that the robot needed 
speech synthesis and a speech recognition system that could acquire new words. These components were available on the 
robot and could therefore be easily integrated. 
\item {\bfshape Flexible scripts.} The Talking Heads experiments used a single very streamlined interaction scenario whereby
the robots played a Guessing Game with a clear turn-taking script. The \textsc{aibo} experiments targeted different 
kinds of interactions with humans and therefore the robot needed not only a way to represent and execute these 
different scripts but also a much more flexible way to move from one step to another within a script and to move 
between scripts. Moreover the scripts had to be integrated intimately with the physical behaviours that were 
steering the robot itself, which could in many cases not be overridden or directly controlled. 
\end{enumerate}

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.50\textwidth]{chap10/figs/look-ball.pdf}
\includegraphics[width=.45\textwidth]{chap10/figs/views-ball.pdf}}
\caption{\label{f:look-ball}Left: Fr\'{e}d\'{e}ric Kaplan interacts with the \textsc{aibo} robot in an experiment for the social learning of lexicons.
Right: View of the ball through the eye of the robot. As the same object is now seen from many different viewpoints it 
becomes much more challenging to consistently recognise it.}
\end{figure}
The first \textsc{aibo} experiments did not involve a population of agents but rather a single agent that was interacting directly with a human
experimenter. The agent had to learn the lexicon of the human through social 
interaction within the context of situated language games. The 2001
paper \dblenquote{\textsc{aibo}'s first words} by Steels and Kaplan \nocite{Steels:01h} described experiments in which an \textsc{aibo} robot 
learned words about objects in the environment, such as \textit{ball} or \textit{smiley}, as well as words for actions, such as 
\textit{sit down} or \textit{stand up}.

@@

The Perspective Reversal experiment \is{Perspective Reversal} focused on the question how a spatial language could emerge that employed
perspective reversal. It was the first experiment to systematically compare different language strategies, thus 
introducing for the first time the {\scshape comparative method} \is{comparative method} 
that would inform all later language game experiments.

A {\scshape language strategy} \is{language strategy} 
prescribes how some aspect of meaning needs to be conceptualised, expressed, and acquired. 
The Talking Heads experiment used a single strategy: find a discriminating category for the topic and employ a word 
for that category. But languages use a variety of strategies all intermixed. 
For example, speakers might have a strategy for expressing argument structure using cases (as in Latin or German) or 
a strategy for expressing tense and aspect using morphological markers on the verb (as in Spanish). Each strategy gives 
rise to a particular feature of a language and in order to demonstrate why languages exhibit this feature it is therefore
possible to compare a language with the feature to one without the feature by doing an experiment where
agents are endowed with a strategy that generates this feature and another one where they employ a different strategy
that does not. The performance of the resulting system, for example, its average communicative success, the time to reach 
linguistic convergence, the amount of cognitive effort involved, can be measured and thus the adaptive value of one 
strategy (and hence one feature of language) can be compared with another.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap10/figs/tokyo-persp.pdf}}
\caption{\label{fig:tokyo-persp}Luc Steels demonstrates live the Perspective Reversal experiment in Tokyo, April 2005. \textsc{aibo} robots track 
the movement of a ball and 
the robot acting as speaker describes to the hearer an observed movement, possibly taking into account the 
perspective of the hearer. For example, the speaker might say \textit{the ball rolls from my left to your right}.}
\end{figure}

\enlargethispage{1\baselineskip}
The Perspective Reversal experiment worked as follows. A population of \textsc{aibo} robots roam around
freely in an unconstrained in-door environment. As soon as one sees the
ball, it comes to a stop and searches for another robot nearby, which also
looks for the ball and stops when it sees it. Then the human
experimenter pushes the ball with a stick so that it rolls a short
distance (see \figref{fig:tokyo-persp}). The two robots next play a description game. One robot (acting as the \dblenquote{speaker}) 
describes the ball-moving event to the other robot (the \dblenquote{hearer}), going through the same semiotic cycle as the Talking 
Heads agents. The game is a success if the meaning conveyed by the speaker is compatible with the hearer's 
own perception of the scene.

@@

Within this experimental setting three different strategies were tried in a long-term experiment where
a new strategy was injected at certain stages and performance monitored: 
\begin{itemize}
\item In stage \textsc{i}, agents use 
the Discrimination Game strategy and the Naming Game strategy discussed in chapters 4 and 5. A 
lexicon is emerging and there are some successful games (those for which the speaker and the hearer see the situation from 
the same perspective). But the lexicon keeps expanding and success is very limited as games where robots have a different 
perspective on the rolling ball all fail. 
\item In stage \textsc{ii}, agents have an extended 
strategy where agents can geometrically transform a model of the scene, based on their own perspective, into a model of the scene 
as seen from the perspective of the hearer. They can then conceptualise the scene from that perspective. 
Communicative success significantly increases and the 
lexicon starts to decrease. However there is ambiguity because the hearer cannot know whether
the speaker has conceptualised the 
scene from his own viewpoint or performed a perspective reversal, leading to a large number of failed communications 
(more than 50 \%). 
\item In stage \textsc{iii}, a more elaborate strategy is used. The speaker now expresses whether he has performed perspective 
reversal (as in {\itshape to} {\scshape your} {\itshape left} vs. simply \textit{left}). Only with this strategy are agents able to reach high 
communicative success with a stable inventory. 
\end{itemize}

@@

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/associations.pdf}}
\caption{\label{fig:associations}Top: Two lexical associations have become active (association-259 and association-234) to cover the distinctive 
feature-set ([VISIB v-411] [VER v-431]) resulting from discrimination and perception. Bottom: 
The group of words and the meanings match with schemas forming part of a grammatical association that combines the words. }
\end{figure}
An application of these schemas in producing the utterance \textit{(W O) (W U)} is represented in \figref{fig:associations} and shown 
here: 
\begin{verbatim}
127 ++> Speaker: head-40 Topic: 2 Context: 6 5 4 3 1 0
Categorial Perception:
  ([VISIB v-4111] [VER v-431])
  ([VER v-4311] [AREA v-4061])
Lexicon lookup: (Association-259 Association-234)
Syntactic structure:
  (syn-cat-77
    (syn-slot-50 (syn-cat-75 |(W O)|))
    (syn-slot-51 (syr-cat-76 |(W U)|)))
Semantic structure:
  (sem-cat-78
    (sem-slot-50 (sem-cat-75 (VER v-431)))
    (sem-slot-51 (sem-cat-76 (VISIB v-411))))
Meaning:
  ([VISIB v-4111] [VER v-4311)
Expression: ((W 0) (W U))
\end{verbatim}

@@

But first we need a clearer idea on what case grammar is.
In the sentence \textit{John gave Mary a book}, John plays the role of agent, the giver 
of the book, Mary that of recipient or 
beneficiary and the book is the \dblenquote{patient} or object of the action. Languages differ in terms of the strategies that they 
use to express these participant roles. Some of them, like German, have an elaborate system of morphological 
markers attached to constituents of the noun phrase. These markers signal that the 
phrase is nominative, accusative, dative, etc., and these cases then map in complex ways 
onto semantic roles such as agent and patient.
Other languages, like English, use a strategy based on word order and prepositions. For example, 
\textit{Mary gave John a book} contains the same participants as \textit{John gave Mary a book} but now Mary is the 
agent and John the recipient. Still other languages, like Japanese, use a strategy based on particles following 
the noun, as in the Japanese sentence \textit{Tanaka-san wa Tokyo de o-to-san ni atta} 
meaning \smplenquote{Tanaka met his father in Tokyo}. \textit{Tanaka-san} is the name of a person, \textit{oto-san} means \smplenquote{father} and \textit{atta} 
means \smplenquote{met}. The markers are \textit{wa} for the topic, \textit{de} for location and \textit{ni} for another sense of location which can 
also function as possessive. The case grammar experiments focused on modelling the latter strategy.

It used the comparative method again to compare three progressively more complex strategies: 
\begin{enumerate}
\item {\bfshape  Lexical strategy}. The experiment starts from a purely 
lexical language that is predefined in all agents. The lexicon contained words for expressing properties of 
objects (such as \textit{smooth}, \textit{red}, etc.)
as well as actions involving objects (such as \textit{move-away-from}, \textit{move-towards} or \textit{push}). English words
were chosen to make the experiment easier to follow. Agents form
multi-word utterances using straightforward lexicon lookup, producing utterances
like \textit{move-towards red jill} (Jill moves towards the red object) or \textit{push 
jack jill green} (Jack pushes the green block to Jill). There is no syntax at all. Agents can nevertheless already 
achieve communicative success when they use the real world as a way to figure out the participant roles, even though the 
utterance does not provide any information. For example, to detect whether Jack pushes the green block to Jill or Jill 
pushes the green block to Jack.

\item {\bfshape  Specific marker strategy}. The second strategy introduces markers in the form of particles added to words referring 
to objects, in order to specify what argument the object fills in the predicate describing the 
event. For example, there could be two markers: \textit{pi} 
and \textit{pa} specifying that the referent plays the role respectively of giver and gift in a give event, as in
\textit{pi jill pa block give}. The markers are entirely specific for each event in the sense that there is no generalisation 
yet to \textit{agent}, \textit{patient} or other more abstract semantic roles. This already yields a more effective 
communication system because the complexity of seeking an interpretation matching with the world model is heavily
reduced and residual ambiguity (i.e. there is still more than one way in which the 
meaning matches with the world model) is also avoided.

@@

\subsubsection{Object identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and \dblenquote{latches onto} regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or \dblenquote{anchors} that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\textsc{cogar} represents syntactic and semantic structures as frames with units, slots and 
constraints on the slots. The frames are in many ways similar to the feature structures used in feature-structure 
based grammars such as \textsc{hpsg}, although there are many important differences. For example, each unit has a 
unique name so that it becomes possible to refer back to units in defining constraints. \textsc{cogar} also used 
logic variables (denoted by a symbol with a question mark in front) as opposed to path descriptions, which 
lead to the adoption of techniques from logic programming. 
The syntactic and semantic structure of a particular utterance are coupled together as a coupled feature structure 
by making the unit-names shared. Often both are combined when the coupled feature structure needs to be displayed
as in this example below. 
\begin{verbatim}
((unit-15
   (scope utterance)
   (subunits (unit-17 unit-18 unit-16))
   (ordering (unit-17 unit-18 unit-16)))
 (unit-17
   (form (smooth))
   (meaning ((smooth ?source-object1)))
   (referent ?source-object1) 
   (gramcat (takes-marker-1)))
 (unit-18 (form (po)))
 (unit-16 
   (form (move-away-from)) 
   (meaning
     ((move-away-from ?event ?state) 
      (move-away-from-1 ?event ?object)
      (move-away-from-2 ?event ?source-object1)))))
\end{verbatim}
This example could have been constructed, either in parsing or producing, for 
the utterance \textit{po smooth move-away-from}. Syntactic properties such as the form of the word, 
the ordering of the units, or the grammatical categories of a unit are represented explicitly on the syntactic side and 
the meaning and referent are represented on the semantic side. In general, \textsc{cogar} allows the representation of
any kind of information at whatever level of linguistic analysis as part of a feature structure.

@@

Here is an example of a lexical rule that defines the meaning of the word for an event \textit{move-away-from}. The referent is 
the object that is referred to by the word, in this case it is the event itself. 
\clearpage
\begin{verbatim}
((?unit 
   (meaning
      (== (move-away-from ?event ?state)
          (move-away-from-1 ?event ?object)
          (move-away-from-2 ?event ?source)))
   (referent ?event)) 
<-->
((?unit
   (form (== move-away-from)))))
\end{verbatim}
Next, here is an example of a rule for the marker \textit{po}. On the semantic side it ensures that the referent of one unit 
whose referent is ?source is equal to the object filling the argument move-away-from-2 in a move-away-from event. 
On the syntactic side, it triggers when there is a word that has the grammatical category \textit{takes-marker-1} and 
is preceded by the marker \textit{po}. (\textless\textless~stands for \smplenquote{precedes}) 
\begin{verbatim}
((?predicate 
   (meaning
    (== (move-away-from ?evnt)
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source))))
 (?argument
   (referent ?source))) 
<---> 
((?argument
   (gramcat (== takes-marker-1)))
 (?marker-unit
   (form (== po))) 
(<< ?marker-unit ?argument))))
\end{verbatim}

There are additional rules that specify which word takes which marker, by assigning them to a particular 
grammatical category. For example, the word \textit{smooth} has the category \textit{takes-marker-1}: 
\begin{verbatim}
((?unit
   (gramcat (== takes-marker-1)) 
   (form (== smooth))))
<---> 
((?unit (form (== smooth))))
\end{verbatim}

The marker \textit{po} in the above example is entirely specific for one argument of the move-away-from predicate, as would be the 
case when the second strategy is enacted. To implement the third strategy 
(to achieve more abstract semantic roles) requires that the markers become generic, which implies that 
we get additional rules that recategorise the arguments of a predicate in terms of 
semantic roles like: agent, patient, source, etc. Here is an example of a rule that 
has this effect. role-1 is an example of a semantic role created by the agent as an abstraction of the 
move-away-from-2 argument. 
\begin{verbatim}
((?predicate
   (meaning
    (== (move-away-from ?evnt) 
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source)))))
<---> 
 ((?predicate
    (meaning
      (== (move-away-from ?evnt)
          (move-away-from-1 ?evnt ?patient)
          (move-away-from-2 ?evnt ?source)))) 
    (expanded-meaning (== (role-1 ?evnt ?source))))
\end{verbatim}
And the rules that click everything together now operate over semantic roles as shown in the following example: 
\begin{verbatim}
 ((?predicate
   (expanded-meaning (== (role-1 ?event ?source))))
  (?unit (referent ?source))) 
<---> 
 ((?unit (gramcat (== takes-marker-1)))
  (?marker-unit (form (== po))) 
 (<< ?marker-unit ?unit))
\end{verbatim}
Recall that all rules must be reversible and that the conditional part must match and the concluding part must unify.
